Assessment and optimization of respiratory syncytial virus prophylaxis regimens in Connecticut, \Sexpr{formatYear(startDate)}-\Sexpr{formatYear(endDate)}

### 2019-02-28 Paper/RSVOptimization.rnw
Typography

Penalized basis splines (``P-splines'') provide a smoothing mechanism with several qualities desirable in analysis of experimental data, including computational compactness and efficiency, and well-understood statistical behavior. Most importantly for our RSV analysis, the parameters of a P-spline approximation have known statistical distributions, which allows interval estimation of arbitrary outcome measures.\cite{Eilers:1996kz}

A simple example of this is our estimation of RSV season onset. We begin the process by defining season onset to occur when cumulative incidence rises above \Sexpr{seasonThreshold * 100}\% of total cases, and obtaining a P-spline approximation to our data (using the \texttt{mgcv} R package). This approximation is characterized by a matrix of coefficients, each of which is normally distributed with a known mean and variance.

@@

More interestingly, the same approach can be used to estimate time-series outcomes. We use this in our computation of preventable cases of RSV. We begin by defining a preventable case of RSV to be a case of RSV that occurs while prophylaxis (under a specific set of prophylaxis guidelines) offers protection; that is, a case that would have been prevented by administration of prophylaxis under those guidelines, if the prophylaxis were 100\% effective. As before, we obtain a P-spline approximation to our data.

We then sample P-spline coefficients from their distributions to obtain a specific approximation to our data, and compute the corresponding time series of preventable RSV cases. Repeating the coefficient-sampling process yields time series of preventable RSV cases sampled from their (again unknown) distribution.

@@

Here we follow a similar process, wherein we sample coefficients from their distributions, use the resultant P-spline to calculate the all-case time series as well as the preventable-case time series, from which we then calculate the preventable fraction; the end result is that we sample the preventable fraction from its (unknown) distribution.

>>>>

Penalized basis splines (``$P$-splines'') provide a smoothing mechanism with several qualities desirable in analysis of experimental data, including computational compactness and efficiency, and well-understood statistical behavior. Most importantly for our RSV analysis, the parameters of a $P$-spline approximation have known statistical distributions, which allows interval estimation of arbitrary outcome measures.\cite{Eilers:1996kz}

A simple example of this is our estimation of RSV season onset. We begin the process by defining season onset to occur when cumulative incidence rises above \Sexpr{seasonThreshold * 100}\% of total cases, and obtaining a $P$-spline approximation to our data (using the \texttt{mgcv} R package). This approximation is characterized by a matrix of coefficients, each of which is normally distributed with a known mean and variance.

@@

More interestingly, the same approach can be used to estimate time-series outcomes. We use this in our computation of preventable cases of RSV. We begin by defining a preventable case of RSV to be a case of RSV that occurs while prophylaxis (under a specific set of prophylaxis guidelines) offers protection; that is, a case that would have been prevented by administration of prophylaxis under those guidelines, if the prophylaxis were 100\% effective. As before, we obtain a $P$-spline approximation to our data.

We then sample $P$-spline coefficients from their distributions to obtain a specific approximation to our data, and compute the corresponding time series of preventable RSV cases. Repeating the coefficient-sampling process yields time series of preventable RSV cases sampled from their (again unknown) distribution.

@@

Here we follow a similar process, wherein we sample coefficients from their distributions, use the resultant $P$-spline to calculate the all-case time series as well as the preventable-case time series, from which we then calculate the preventable fraction; the end result is that we sample the preventable fraction from its (unknown) distribution.
